---
title: "Population Structure of Ouachita Bass"
author: "Joe Gunn"
date: "2022-08-14"
output: html_document
---

# Project: Smallmouth, Neosho, Ouachita, and Little River Hybridization (SNOLH)
<font size="+1">Investigating hybridization and population structure among and within the Smallmouth Bass (<i>Micropterus dolomieu</i>), the newly elevated Neosho Bass (<i>M. velox</i>), and two other potentially distinct species in the Ouachita River Basin (the Ouachita and Little River Basses)</font>

## Specific Aim: Hierarchical structure in Ouachita Bass
<font size="+1">For this aim, we are using Bayesian inference with STRUCTURE and NEWHYBRIDS to assess population structure in the Ouachita River Basin (Ouachita Bass).

### Libraries needed for analysis
```{r}
library(tidyverse)
library(cowplot)
library(readxl)
library(writexl)
library(genepopedit)
library(pophelper)
library(parallelnewhybrid)
```

### STEP 1: Data preparation

#### 1a: Load in full raw dataset ('full_data'), metadata ('metadata'), and genotype data ('genotype_data')

##### 1a.1: Run the Rmd chunk below to load in all data.

##### Load data:
```{r}
load("data/raw_data/snolh_all_data.rda")
load("data/raw_data/snolh_metadata.rda")
load("data/raw_data/snolh_genotype_data.rda")
```

#### 1b: Filter metadata and SNP genotype data 
In this step, we are filtering the metadata and SNP genotype data to only include samples of interest for analysis of population structure of Ouachita Bass. For this analysis, these data will include Ouachita Bass and only SNP loci that were found to be diagnostic for Ouachita Bass.

##### 1b.1. Filter metadata; run the Rmd chunk below:

##### Filter metadata to only include Ouachita Bass
```{r}
ouachita_metadata <- ouachita_little_metadata %>%
  filter(Putative_Taxon != "Spotted") %>%
  filter(Putative_Taxon != "Smallmouth_Northern") %>%
  filter(Putative_Taxon != "Smallmouth_TN_Native") %>%
  filter(Putative_Taxon != "Smallmouth_TNLakeStrain") %>%
  filter(Putative_Taxon != "Neosho")
  filter(Putative_Taxon != "Ouachita_LittleRiver") # Only Ouachita Bass
```

##### 1b.2. Filter SNP genotype data; run the Rmd chunk below:

##### Filter out genotype data into diagnostic SNP groups for Ouachita Bass and Little River Bass
```{r}
ouachita_snps <- full_data %>%
  filter(Putative_Taxon != "Spotted") %>%
  filter(Putative_Taxon != "Smallmouth_Northern") %>%
  filter(Putative_Taxon != "Smallmouth_TN_Native") %>%
  filter(Putative_Taxon != "Smallmouth_TNLakeStrain") %>%
  filter(Putative_Taxon != "Neosho") %>%
  filter(Putative_Taxon != "Ouachita_LittleRiver") %>%
  select(matches("OUOU"))

# N(sample) = 159
# N(loci) = 46
```

#### 1c: Combine metadata and snp datasets for combinations of interest (see above); Run the Rmd chunk below to combine datasets.

##### Combine metadata and snp datasets and generate excel/Rda files: files are stored in the directory `data/processed_genotype_data/`
```{r}
# Ouachita Bass only:
ouachita <- cbind(ouachita_metadata, ouachita_snps)

# Save combined datasets .Rda files to be loaded in downstream analysis
save(ouachita, file = "data/processed_genotype_data/ouachita.rda")

# From this point, we are exporting processed genotype data to excel so that the datasets can be converted easily into genepop format. Subsequent, processed genepop files are saved in the directory

## Generate excel files to hold genotype data and to prepare:
write_xlsx(ouachita, path = "data/processed_genotype_data/ouachita.xlsx")
```

#### 1d: Generate STRUCTURE formatted files from genepop format
In this step, we are reading in, checking, and converting genepop files for combined datasets into STRUCTURE formatted files. There are quite a few tricky and very specific formatting needs for the genepop text file before genepopedit will convert successfully. Follow Step 3a below closely to create a proper genepop format. Use Ctrl + F to find and replace the necessary things.

##### IMPORTANT NOTE ON DATA: genepopedit requires that sample names be in the format 'popname_01', 'popname_02', etc. where the 'popname' is a signifier for the population of origin for the sample, and the '01', etc. are numerical designations for the samples. 'popname' and the numeric must be separated by a "_", and there can only be a single "_". Otherwise, genepopedit will not read the sample's population of origin correctly. In the case of this project, we are clumping samples together by stream for STRUCTURE analysis to see if there may be substructure mapped by stream. Many sample IDs in our dataset were not in the required format, so we added an additional column to the metadata set and associated processed data sets called "structure_ID", which gives a unique sample name in the proper structure format. This column can then be linked back to any other column in downstream analyses if we need to know which specific samples the results came from.

#### 1d.1. Follow the guide for the library 'genepopedit' (Stanley et al. [year]) to prepare genepop files before converting them to STRUCTURE format.

###### 1d.1.1. All "," in the genepop text files were manually replaced with " ,  "

###### 1d.1.2. All "tabs" between six-digit alleles were were manually replaced with " " (a single space).

###### 1d.1.3. The top row designating the dataset combination was manually omitted.

###### 1d.1.4. Any additional space at the bottom of the text file was removed.

##### 1d.2. Generate STRUCTURE files from genepop files
Here it is important to not include the 'popgroup = ' unless it is explicitly necessary. Assigning popgroup to a dataframe with population names will cause the STRUCTURE input file to have word strings in the popdata column. Our experience is that STRUCTURE expects an integer here rather than a string, so best not to use popgroup.

##### Convert genepop to STRUCTURE format: files are stored in the directory `data/structure_data/`
```{r}
# Generate structure file for Ouachita Bass only
genepop_structure("data/genepop_data/ouachita_genepop.txt", 
                  locusnames = TRUE, 
                  #popgroup = "data/genepop_data/ouachita_pops.csv",
                  path = "data/structure_data/input_data/ouachita_structure.txt")

# N(pops) = 4
```

#### 1e: Generate batch list files and shell scripts for running STRUCTURE in parallel
In this step, we are creating batch lists of command line code to run STRUCTURE analyses in parallel. We are making a single batch list for each of our data combinations (generated above; 'spb_other', 'smb_ih', 'ih', 'ouachita_little', 'ouachita', 'little'). Each batch list will contain a separate line of code to run a single replicate at an a priori determined number of populations (<i>K</i>, listed in the chunk above for each dataset). We will run each analysis in 10 replicates at each <i>K</i>, e.g.:

structure -K 1 -m mainparams -e extraparams -i structure_input.txt -o structure_output_1_1
structure -K 1 -m mainparams -e extraparams -i structure_input.txt -o structure_output_1_2
structure -K 1 -m mainparams -e extraparams -i structure_input.txt -o structure_output_1_3
structure -K 1 -m mainparams -e extraparams -i structure_input.txt -o structure_output_1_4
structure -K 1 -m mainparams -e extraparams -i structure_input.txt -o structure_output_1_5
.
.
.
structure -K 1 -m mainparams -e extraparams -i structure_input.txt -o structure_output_1_10
structure -K 2 -m mainparams -e extraparams -i structure_input.txt -o structure_output_2_1
structure -K 2 -m mainparams -e extraparams -i structure_input.txt -o structure_output_2_2
.
.
.
where, in the output name, the first number represents the K value for the run, and the last number represents the replicate.

##### 1e.1. Generate batch list file; run the Rmd chunk below:
The Rmd chunk below generates a list of commands that contains the full cluster paths (/home/data/...) for necessary files.

##### Generate batch commands: `code/structure_code/batch_cmd_lists/ouachita_structure_batch_cmd_list.txt`
```{r}
# N(pops) = 4
nk <- data.frame(c(1:4))
nreps <- data.frame(c(1:10))

cat("", file="code/structure_code/batch_cmd_lists/ouachita_structure_batch_cmd_list.txt")

# Run loop to create file for storing commands
for(ii in 1:nrow(nk)) {

  for(aa in 1:nrow(nreps)) {
    
    structure_call <- paste("structure ")

    param_files <- paste(" -m /home/jcg5g9/data/SNOLH_Genomics/structure_analysis/data/structure_data/param_files/mainparams_ouachita -e /home/jcg5g9/data/SNOLH_Genomics/structure_analysis/data/structure_data/param_files/extraparams")

    input <- paste(" -i /home/jcg5g9/data/SNOLH_Genomics/structure_analysis/data/structure_data/input_data/ouachita_structure.txt")
  
    output <- paste(" -o /home/jcg5g9/data/SNOLH_Genomics/structure_analysis/data/structure_data/output_data/ouachita_structure_output/")
  
    cat(paste(structure_call, "-K ", nk[ii,], param_files, input, output,
            strsplit("ouachita_structure.txt", "_structure.txt", fixed=TRUE)[[1]][1], "_", nk[ii,], "_", nreps[aa,], sep=""),
      "\n", 
      file=paste("code/structure_code/batch_cmd_lists/ouachita_structure_batch_cmd_list.txt"), # name file to store list
      append=TRUE)
  }
  
}
```

##### 1e.2. Generate shell script; run the Rmd chunk below:

##### IMPORTANT NOTE: This shell script was generated on the server directly, and thus the Rmd chunk below does not need to be run to generate the file. This is purely to keep a record of each script file. 

##### Generate batch commands: `code/structure_code/shell_scripts/ouachita_structure.sh`
```{r}
#!/bin/bash
#-------------------------------------------------------------------------------
#  SBATCH CONFIG
#-------------------------------------------------------------------------------
## resources
#SBATCH --partition Lewis
#SBATCH --nodes=1
#SBATCH --ntasks=1  # used for MP#SBATCH -e error_%A_%a.err # Standard errorI codes, otherwise leav$
##SBATCH --cpus-per-task=12  # cores per task
#SBATCH --mem-per-cpu=16G  # memory per core (default is 1GB/core)
#SBATCH --time 1-00:00  # days-hours:minutes
#SBATCH --qos=normal
#SBATCH --array=23-220

## labels and outputs
#SBATCH --job-name=snolh_structure_jgunn
#
#SBATCH -o test_%A_%a.out # Standard output
#SBATCH -e error_%A_%a.err # Standard error

## notifications
#SBATCH --mail-user=jcg5g9@mail.missouri.edu  # email address for notifications
#SBATCH --mail-type=END,FAIL  # which type of notifications to send
#-------------------------------------------------------------------------------

#echo "### Starting at: $(date) ###"


# load packages
#module load rss/rss-2020
#module load structure/structure-2.3.4

#COMMANDA=`head -n ${SLURM_ARRAY_TASK_ID} ../batch_cmd_lists/ouachita_structure_batch_cmd_list.txt | tail -n 1`
#eval $COMMANDA


#echo "### Ending at: $(date) ###"
```

#### 1f: Prepare mainparam and extraparam files for STRUCTURE.
In this step, we are preparing the mainparam and extraparam input files for STRUCTURE so that they are unique to each analysis we are running.

##### 1f.1. Edit base mainparam STRUCTURE file (downloaded with STRUCTURE program) and generate a separate, unique mainparam file.

<b>mainparam file:</b> <br>
`data/structure_data/param_files/mainparams_ouachita` <br> 

###### 1f.1.1. Edit "maxpops" value to reflect the number of populations designated in Step 3b above. These values should be the following:

<b>Ouachita Bass populations only:</b> 4 <br>

###### 1f.1.2. Set the number of burn-in and MCMC iterations to run; these are the same for each analysis:

<b>Burn-in runs:</b> 500,000 <br>
<b>MCMC runs:</b> 1,000,000 <br>

###### 1f.1.3. Set the number of individuals. These values should be the following:

<b>Ouachita Bass populations only:</b> 159 <br>

###### 1f.1.4. Set the number of loci. These values should be the following:

<b>Ouachita Bass populations only:</b> 46 <br>

###### 1f.1.5. Set 'ONEROWPERIND' to '0'

###### 1f.1.6. Set 'LABEL' to '1'

###### 1f.1.7. Set 'POPDATA' to '1'

###### 1f.1.8. Set 'POPFLAG' to '0'

##### 1f.2. Edit base extraparam STRUCTURE file (downloaded with STRUCTURE program).
We did not change any of the default settings in the extraparams file (most importantly, we used the default admixture model), so we only used a single extraparams file.

<b>extraparam file:</b> <br>
`data/structure_data/param_files/extraparams` <br> 

### STEP 2: population Structure Analysis

#### 2a: Run STRUCTURE analysis using the input data generated in STEP 1. Navigate to `code/structure_code/shell_scripts/` Be sure that all relative and full paths to all input files and output destination directories are set up properly (ideally, this is already done within this GitHub repo). This command line code assumes capability to run the code using SLURM or a SLURM-like cluster scheduling software.

sbatch ouachita_structure.sh

#### 2b: Structure output files are generated and stored here: `data/structure_data/output_data/`.

#### 2c: Compress output directory into a zip file compatible with Structure Selector (Li and Liu 2017) or Structure Harvester (Earl and vonHoldt 2011) online.

#### 2d: Submit zip directory to Structure Selector or Structure Harvester to extract summary results

#### 2e: Copy detlaK table (Evanno et al. YEAR) and save as excel file here: `data/structure_data/summary_data`

#### 2f: Copy puechmaille table (Puechmaille et al. 2016.) and save as excel file here:
`data/structure_data/summary_data`

#### 2g: Visualize STRUCTURE runs for all data combinations

##### 2g.1. Convert STRUCTURE files into aligned Q files compatible with analysis in the program CLUMPP (Jakobbson and Rosenberg 2007); run the rmd chunk below.

##### Convert STRUCTURE files to aligned Q files for CLUMPP: 
```{r}
# Get a list of structure files for each data combination
ouachita_sfiles <- list.files("data/structure_data/output_data/ouachita_structure_output/", full.names = T)

# Extract q value information (ancestry proportions) from each run for each K value for each individual from the STRUCTURE output files in the directories listed above
ouachita_Q <- readQ(ouachita_sfiles)

# Tabulate information from the q lists
ouachita_tab <- tabulateQ(ouachita_Q)

# Summarize information from tabultions above
ouachita_summary <- summariseQ(ouachita_tab)

# Extract deltaK and associated summary information using Evanno method
ouachita_evanno <- evannoMethodStructure(ouachita_summary, returnplot = F) 

# Align replicate runs for each K to correct label switching
ouachita_align <- alignK(ouachita_Q)
```

##### 2g.2. Export CLUMPP compatbile files for CLUMPP analysis; run the rmd chunk below to export CLUMPP associated files for each K.

<b>We used the following parameters: </b> <br>

Large-K-Greedy algorithm (paramrep = 3) <br>
10000 replicates <br>

### IMPORTANT NOTE: This step only needs to be run ONCE to generate files for CLUMPP. Once you have run this chunk, move on to Step 2g.3. Uncomment each line to run this code.

##### Export CLUMPP files:
```{r}
clumppExport(ouachita_align, 
            parammode = 3,
            paramrep = 10000, 
            exportpath = "data/structure_data/clumpp_data/ouachita_clumpp")
```

The code above generates a .txt file (pop_K#-combined.txt) with combined cluster ancestry proportions for each individual at each K and stores it in a separate directory labeled "pop_K#", where '#' is the corresponding K value. Additionally, the code generates an accompanying "paramfile" that is input for CLUMPP, which contains information on parameters to generate merged datasets across our 10 replicates at each K.

##### 2g.3. Generate merged Q files using CLUMPP.
In this step, we used the software program CLUMPP (Jakobbson and Rosenberg 2007) to merge cluster ancestry proportion inferences for all replicate runs of each K across individuals. We downloaded the Linux 64-bit build of CLUMPP v.1.1.2 and installed it on our computing cluster. We then executed the program for each analysis by sequentially copying the associated paramfiles generated in the clumppExport function (see section above) to the CLUMPP home directory. All output files were then moved to the corresponding structure directory.

###### 2g.3.1. Downolad CLUMPP and install on cluster; navigate to the desired directory: `code`. Run the command: `wget https://rosenberglab.stanford.edu/software/CLUMPP_Linux64.1.1.2.tar.gz`

###### 2g.3.2. Run the command: `gunzip CLUMPP_Linux64.1.1.2.tar.gz`

###### 2g.3.3. Run the command: `tar -xvf CLUMPP_Linux64.1.1.2.tar`

###### 2g.3.4. For each analysis separately, copy the 'pop_K#-combined.txt' file and 'paramfile' over to the CLUMPP directory (`code/CLUMPP_Linux64.1.1.2`)

###### 2g.3.5. For each analysis separately, execute CLUMPP on the paramfile; run the code `./CLUMPP paramfile`
We are running this code with GREEDY OPTION 2, which uses random sampling of replicate runs. 

This code generates three additional files, which we moved back to the corresponding pop_K folder in the clumpp_data directory:

<b>Files generated: </b> <br>

pop_K#-combined-miscfile.txt <br>
pop_K#-combined-merged.txt <br>
pop_K#-combined-aligned.txt <br>

##### 2g.4. Plot STRUCTURE runs at best K value (according to deltak statistic) for each analysis; run the rmd chunk below.

##### STRUCTURE results for Ouachita populations
```{r}
# DeltaK results
ggplot(ouachita_evanno, aes(x=k, y=deltaK)) +
  geom_point() + 
  geom_line() +
  theme_cowplot(theme_set(12)) +
  geom_vline(xintercept = 2, color = "blue") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 15)) +
  theme(axis.title.x = element_text(face = "italic")) +
  labs(x = "K", y = "delta K") + 
  scale_x_continuous("K", labels = as.character(ouachita_evanno$k), breaks = ouachita_evanno$k)

# convert metadata columns to characters 
ouachita <- ouachita %>%
  mutate(Structure_Num = as.character(Structure_Num),
         River = as.character(River))

ouachita_k2 <- readQ("data/structure_data/clumpp_data/ouachita_clumpp/pop_K2/pop_K2-combined-merged.txt")

# Visualize Structure
plotQ(ouachita_k2,
      grplab = ouachita[,c(12,6)],
      showindlab = FALSE, 
      ordergrp = F,
      selgrp = "River",
      showgrplab = T,
      showlegend = F, 
      showsp = F, 
      showdiv = T, 
      divsize = 2, 
      divtype = 1,
      divcol = "black",  
      grplabsize = 4, 
      grplabangle = 45,
      grplabjust = 0.1,
      legendkeysize = 15, 
      linealpha = 0,
      legendtextsize = 10, 
      linesize = 0.05, 
      pointsize = 4, 
      barbordercolour = "black",
      barbordersize = 0.1, 
      clustercol = c("deepskyblue","deeppink2"),
      exportplot = T,
      imgtype = "pdf",
      height = 10, 
      width = 100,
      outputfilename = "ouachita_k2",
      exportpath = "figures/structure_figures")
```

### STEP 3: Hybridization analysis with NEWHYBRIDS

See `snolh_structure_analysis.Rmd` (Line XXX) for programs needed for analysis.


